### 2025-01-22 analysis/preprocessing.qmd

```{r}
#| warning: false
#| echo: false
#| eval: true
# Set parameters and libraries
export_decision <- TRUE
verbose <- FALSE
require(here)
library(stringr)
library(tidyverse)
library(ggrepel)
library(ez)
library(ggdist)
```

@@

```{r}
#| echo: false
#| eval: true
# this entry no longer appears due to exclusion:
# R_studies$feature_n[R_studies$feature_n==" 548; after reduction, 139 for PCA and 276 for ReliefF   "] <- 548
# R_studies$feature_n[str_detect(R_studies$feature_n,'548 dimensions. Pos')]<-548
R_studies$feature_n[str_detect(R_studies$feature_n,'pre_fitting')] <- 21
R_studies$feature_n[str_detect(R_studies$feature_n,'but 15 reported')]<-15 #resolved from the paper
R_studies$feature_n[str_detect(R_studies$feature_n,'150 PCA features')]<-3000 #discuss
R_studies$feature_n[str_detect(R_studies$feature_n,' 50 PCA features')]<-  499 #resolved
R_studies$feature_n[str_detect(R_studies$feature_n,'60 handcrafted and')]<-60 # this should be 14400+60! discuss
R_studies$feature_n[str_detect(R_studies$feature_n,'before_selection = 45')]<-45 #resolved
R_studies$feature_n[str_detect(R_studies$feature_n,'model 4 = 388')]<-388 #resolved
R_studies$feature_n[str_detect(R_studies$feature_n,'6670 in MediaEval')]<-6670 #discuss
R_studies$feature_n[str_detect(R_studies$feature_n,'557 before feature')]<-557 #resolved
# this seems to refer to yang2021an, which is handled above with "6670 in mediaeval"
# R_studies$feature_n[str_detect(R_studies$feature_n,'not specified')]<-NA
# 2025-01-15
# beveridge2018po. This one is worth discussing as the two studies might be worth differentiating. 
# Using higher value for now. DISCUSS
R_studies$feature_n[str_detect(R_studies$feature_n,'study 1: 26, study 2: 28')] <- 28
R_studies$feature_n<-as.numeric(R_studies$feature_n)

@@

#table(C_studies$feature_n)
C_studies$feature_n[str_detect(C_studies$feature_n,'eight different ')]<-600 # arbitrary!
C_studies$feature_n[str_detect(C_studies$feature_n,'126, retained 97')]<-126 
C_studies$feature_n[str_detect(C_studies$feature_n,'1702; best model uses 100 after reduction')]<-1702 #  arbitrary!
# these seem to refer to the excluded bai studies
# C_studies$feature_n[str_detect(C_studies$feature_n,'548; after reduction, ')] <- 548 
#C_studies$feature_n[str_detect(C_studies$feature_n,'548 dimensions')] <- 548 
# C_studies$feature_n[str_detect(C_studies$feature_n,'548. Post-reduction')] <- 548 
C_studies$feature_n[str_detect(C_studies$feature_n,'8; 3 after')]<-8 
C_studies$feature_n[str_detect(C_studies$feature_n,'between 9 and 10')]<-10 
C_studies$feature_n[str_detect(C_studies$feature_n,'summarize feature')]<-600 # arbitrary
#table(C_studies$feature_n)

>>>>

```{r}
#| warning: false
#| echo: false
#| eval: true
# Set parameters and libraries
export_decision <- FALSE
verbose <- FALSE
require(here)
library(stringr)
library(tidyverse)
library(ggrepel)
library(ez)
library(ggdist)
```

@@

```{r}
#| echo: false
#| eval: true
# this entry no longer appears due to exclusion:
# R_studies$feature_n[R_studies$feature_n==" 548; after reduction, 139 for PCA and 276 for ReliefF   "] <- 548
# R_studies$feature_n[str_detect(R_studies$feature_n,'548 dimensions. Pos')]<-548
R_studies$feature_n[str_detect(R_studies$feature_n,'pre_fitting')] <- 21
R_studies$feature_n[str_detect(R_studies$feature_n,'but 15 reported')]<-15 #resolved from the paper
R_studies$feature_n[str_detect(R_studies$feature_n,'150 PCA features')]<-3000 #resolved
R_studies$feature_n[str_detect(R_studies$feature_n,' 50 PCA features')]<-  499 #resolved
R_studies$feature_n[str_detect(R_studies$feature_n,'60 handcrafted and')]<-14400+60 # this should be 14400+60! resolved
R_studies$feature_n[str_detect(R_studies$feature_n,'before_selection = 45')]<-45 #resolved
R_studies$feature_n[str_detect(R_studies$feature_n,'model 4 = 388')]<-388 #resolved
# yang2021an appears there are nine unique features based on table 1
R_studies$feature_n[str_detect(R_studies$feature_n,'6669 in MediaEval')]<-6669 # resolved. 9 reported in table 1, but likely used full mediaeval feature set. 
R_studies$feature_n[str_detect(R_studies$feature_n,'557 before feature')]<-557 #resolved
# this seems to refer to yang2021an, which is handled above with "6669 in mediaeval"
# R_studies$feature_n[str_detect(R_studies$feature_n,'not specified')]<-NA
# 2025-01-15
R_studies$feature_n[str_detect(R_studies$feature_n,'study 1: 26, study 2: 28')] <- 54 # resolved used audio and midi features. study 2 was external validation
R_studies$feature_n<-as.numeric(R_studies$feature_n)

@@

#table(C_studies$feature_n)
C_studies$feature_n[str_detect(C_studies$feature_n,'eight different ')]<-21+12+4+14+39+1+1+6 # estimated from feature subsection 4.1.2 (encoded in reverse order from h to a)
C_studies$feature_n[str_detect(C_studies$feature_n,'126, retained 97')]<-126 
C_studies$feature_n[str_detect(C_studies$feature_n,'1702; best model uses 100 after reduction')]<-1702 #  resolved
# these seem to refer to the excluded bai studies
# C_studies$feature_n[str_detect(C_studies$feature_n,'548; after reduction, ')] <- 548 
#C_studies$feature_n[str_detect(C_studies$feature_n,'548 dimensions')] <- 548 
# C_studies$feature_n[str_detect(C_studies$feature_n,'548. Post-reduction')] <- 548 
C_studies$feature_n[str_detect(C_studies$feature_n,'8; 3 after')]<-8 
C_studies$feature_n[str_detect(C_studies$feature_n,'between 9 and 10')]<-10 
C_studies$feature_n[str_detect(C_studies$feature_n,'summarize feature')]<-2+2+2 # estimated from 3.2.1 to 3.2.3
#table(C_studies$feature_n)

### 2025-01-22 manuscript/_figure1.qmd

digraph sub {
  node [fontname = Helvetica, fontsize = 14, shape = box, width = 4]
  A [label = "Records identified from:
                Web of Science (n = 142)
                Scopus (n = 227)
                OpenALEX (n = 278)" ]
  B [label = "Records removed before screening: 
                Duplicates entries/errors (n = 94)" ]
  C [label = "Records after duplicates removed (n = 553)"]
  D [label = "Records screened (n = 553)"]
  E [label = "Records removed: 
                Irrelevant titles/themes (n = 338)
                Abstract screening (n = 95)
                Discussion (n = 24)"]
  F [label = "Full text articles assessed (n = 96)"]
  G [label = "Studies included in final review (n = 46)"]
  H [label = "Full text articles excluded
             (n = 50)"]
  I [label = "Studies excluded in final review (n = 10)"]
  J [label = "Final selection (n = 36)"]
  A->B
  A->C
  C->D
  D->E
  D->F
  F->H 
  F->G
  G->I
  G->J

>>>>

digraph sub {
  node [fontname = Helvetica, fontsize = 14, shape = box, width = 4]
  A [label = "Records identified from:
                Web of Science (n = 142)
                Scopus (n = 227)
                OpenALEX (n = 278)" ]
  B [label = "Records removed before screening: 
                Duplicates entries/errors (n = 94)" ]
  C [label = "Records after duplicates removed (n = 553)"]
  D [label = "Records screened (n = 553)"]
  E [label = "Records removed: 
                Irrelevant titles/themes (n = 338)
                Abstract screening (n = 95)
                Discussion (n = 24)"]
  F [label = "Full text articles assessed (n = 96)"]
  G [label = "Studies included in final review (n = 46)"]
  H [label = "Full text articles excluded
             (n = 50)"]
  I [label = "Studies excluded in final review (n = 12)"]
  J [label = "Final selection (n = 34)"]
  A->B
  A->C
  C->D
  D->E
  D->F
  F->H 
  F->G
  G->I
  G->J

### 2025-01-22 manuscript/manuscript.qmd

---
title             : "Meta-analysis of regression and classification success of emotion ratings from audio"
shorttitle        : "Meta-analysis of music emotion recognition"

@@

The initial search yielded 553 potential studies after excluding duplicate entries. We interactively screened them for relevance in three stages, resulting in 46 studies that passed our inclusion criteria (music emotion studies using classification or regression methods to predict emotion ratings of music using symbolic or audio features, and containing sufficient detail to convert results to $R^2$ or $MCC$ values see [SI](http://) for a breakdown). After the screening stage, we defined a set of entities to extract characterising (i) music (genre, stimulus number \[N\], duration), (ii) features extracted (number, type, source, defined by [@panda2020audio]), (iii) model type (regression, neural network, SVM, etc.) and outcome measure ($R^2$, *MSE*, *MCC*), (iv) model complexity (i.e., approximate number of features used to predict ratings), and (v) type of model cross-validation.

We converted all regression results into $R^2$ values for valence and arousal and classification results into Matthews correlation coefficient [*MCC*, @chicco2020advantages]. To increase consistency in our analyses, we excluded studies using incompatible features (e.g., spectrograms of audio files [@nag2022]), or dependent variables (e.g., one regression study analyzed valence and arousal together, but not separately [@chin2018]).

@@

The search yielded studies of variable (and occasionally questionable) quality. To mitigate potentially spurious effects resulting from the inclusion of low-quality studies, [we excluded studies lacking sufficient details about stimuli, analyzed features, or model architecture](include%20approx%20#%20for%20each?%20This%20would%20be%20good%20since%20the%20rest%20of%20this%20para%20already%20mentions%20counts.%20If%20this%20take%20a%20long%20time,%20we%20could%20save%20this%20for%20an%20additional%20analysis). Finally, we excluded studies published in journals of questionable relevance/quality, (e.g., *Mathematical Problems in Engineering* ceased publication following 17 retractions published between July and September 2024).

@@

TABLE 1: Summary of data (part of `analysis/preprocessing.qmd`)

|            | Regression                      | Classification            | Total |
|:-----------|:--------------------------------|:--------------------------|:------|
| Study N    | 21                              | 14                        | 36    |
| Model N    | 236                             | 103                       | 339   |
| Techniques | Flexible Discriminants (FD): 58 | 30                        | 86    |
| Techniques | Neural Nets (NN): 70            | 28                        | 98    |
| Techniques | Linear Methods (LM): 62         | 13                        | 75    |
| Techniques | Random Forests (RF): 22         | 14                        | 36    |
| Techniques | KS, Add. & KNN (KS): 24         | 8                         | 32    |
| Feature N  | Min=3, Md=548, Max=6670         | Min=3, Md=397, Max=8904   | NA    |
| Stimulus N | Min=20, Md=330, Max=2486        | Min=124, Md=387, Max=5192 | NA    |

Although the total number of studies meeting the criteria described in the previous section is modest (34 in total), they encompass a large array of models (339 in total) with a relatively even distribution among the three most popular techniques: flexible discriminants, neural nets, and linear methods. The number of features and stimuli within these studies varies significantly, ranging from as few as three features [@battcock2021in] to a maximum of almost 9000 features [@zhang2023mo]. The median number of features differs between regression (548) and classification (397) studies, primarily reflecting the nature of the datasets used in each approach. The number of stimuli is typically around 300-400 (with a median of 330 for regression and 387 for classification), though there is substantial variation, with the extremes from 40 stimuli in @saizclar2022pr to 5192 stimuli in @alvarez2023ri. There are also additional dimensions to consider, such as the type of cross-validation used, the music genres analyzed (whether a single genre, multiple genres, or a mix), the type of journal in which the studies were published, and the source of the extracted features. However, these variables do not lend themselves to a simple summary, so we will revisit them during the interpretation and discussion stages.

@@

See `analysis/analysis.qmd`

@@

| Concept      | Models, obs | $r$ \[95%-CI\]          | $t$   | $p$   | $\tau^2$ | $I^2$ |
|:-------------|:------------|:------------------------|:------|:------|:---------|:------|
| Valence All  | 90,61733    | 0.609 \[0.570-0.650\]   | 23.3  | .0001 | 0.079    | 97.3% |
| Valence Max  | 21,13572    | 0.680 \[0.575-0.762\]   | 9.96  | .0001 | 0.136    | 98.2% |
| *N Features* |             |                         |       |       |          |       |
| \<18 F       | 5,3036      | 0.811 \[0.542-0.823\]  | \-    | \-    | 0.182    | 98.9% |
| 18-260 F     | 11,7042     | 0.584 \[0.534-0.931\]  | \-    | \-    | 0.133    | 97.6% |
| 260+ F       | 5,3494      | 0.710 \[0.542-0.823\]  | \-    | \-    | 0.044    | 97.2% |
| *Techniques* |             |                         |       |       |          |       |
| KS           | 2,2582      | 0.466 \[-0.634-0.942\]  | \-    | \-    | 0.0185   | 95.1% |
| LM           | 7,1438      | 0.797 \[0.614-0.899\]   | \-    | \-    | 0.1370   | 96.4% |
| FD           | 6,4993      | 0.656 \[0.484-0.779\]  | \-    | \-     | 0.0574   | 96.9% |
| NN           | 3,1229      | 0.393 \[-0.355-0.835\] | \-    | \-     | 0.0761   | 97.1% |
| RF           | 3,3330      | 0.750 \[0.292-0.928\]  | \-    | \-     | 0.093    | 98.5% |

The results indicate that valence can generally be predicted with moderately accuracy, with the best model from each of the 24 studies achieving an average correlation of _r_ = 0.658 (95% CI: 0.557-0.740), called "valence Max" in Table 2. However, when considering all models across these studies (n = 120), the overall prediction rate drops significantly to _r_ = 0.567. We argue that this lower correlation is likely due to the inclusion of baseline models reported in these studies, which may not reflect the true success of the task for the purposes of our analysis.

Further analysis of between-study heterogeneity, as indexed by the $\tau^2$ (0.138) and Higgins & Thompson's $I^2$ statistic [@higgins2002quantifying] at 98.2%, reveals substantial heterogeneity. Since $I^2$ is heavily influenced by study size (with larger N leading to lower sampling error), its value may be less insightful in this context. In contrast, $\tau^2$, which is less sensitive to the number of studies and directly linked to the outcome metric (_r_), provides a more reliable measure of heterogeneity in this case. Also, we note that because the overall heterogeneity in the data is high, we are cautious in our interpretation of the publication bias [@van-aertwicherts_2016].

To better understand the effects across studies and the nature of the observed heterogeneity, Figure 2 presents (A) a forest and (B) funnel plot of the random-effects model, based on the best-performing models from all studies. In terms of the forest plot, the range of prediction values (correlations) is broad, spanning from 0.13 to 0.92, with all studies except Koh et al. (2023) demonstrating evidence of positive correlations. A mean estimate of 0.66 is achieved by 15 out of the 24 models. While the confidence intervals are generally narrow due to the large sample sizes in each study, there are exceptions, such as smaller sample sizes in @beveridge2018po (n = 20), @saizclar2022pr (n = 40), and in @griffiths2021am (n = 40). The funnel plot in panel B of Figure 2 shows clustering at the top of the plot (studies with low standard error) and no assumed larger diversity in the correlations when the error rates increase. However, there is no clear asymmetry in the plot, verified by non-significant Egger's test ($\beta$=5.00, CI95% -0.78-10.78, _t_ = 1.70, _p_ = 0.104, @eggersmith_1997). Coming back to the mean of valence correlation of 0.66 by all studies and the possible impact of study heterogeneity on this estimation, we also calculated the correlation without the studies that lie outside the 95% CI for pooled effect. This left 13 studies in the data and resulted in the meta-analytical pooled correlation of 0.649 (CI95% 0.598-0.695). In other words, despite the large variation in the correlations and standard errors across the studies, this variation in itself does not seem to be a significant driver behind the overall effect.

@@

To gain insights into the factors contributing to the wide range of model success, we explored several ways of splitting the data. Table 2 presents two key splits: one based on the number of features used, which we hypothesized might influence model performance, and another based on the modeling techniques employed. In terms of feature sets, we categorized them into three groups: few features (<18), a large number of features (18–260), and massive feature sets (260+). Interestingly, models using a relatively small number of features (<18) performed best (_r_ = 0.881, 95% CI: 0.566–0.???) compared to those utilizing larger feature sets. However, it is worth noting that the models using massive feature sets (260+) also performed well (_r_ = 0.685), achieving better and more consistent results than the overall prediction rate (_r_ = 0.659). This observation is supported by the lowest heterogeneity index for the massive feature set group ($\tau^2$ = 0.044), indicating more consistent results across studies.

When analyzing the studies across the five modeling techniques used. Notably, linear models (LM) and flexible discriminants (FD) were the most common, with 8 and 6 studies, respectively, allowing for more confident interpretations. Linear models achieved the highest prediction rate (_r_ = 0.784, 95% CI: 0.625–0.881), though this may be influenced by the smaller datasets typically used in these studies. These studies also exhibited substantially higher heterogeneity ($\tau^2$ = 0.137) compared to other techniques, where $\tau^2$ values were less than half of this. Flexible discriminants performed at a similar level to the overall model average (_r_ = 0.656, 95% CI: 0.484–0.779). The relatively poor performance of the neural network (NN) models (_r_ = 0.340, 95% CI: -0.097–0.668) is difficult to explain without a deeper examination of the specific model architectures and the stimuli used in these studies.

We also ran the sub-grouping analyses across a combination of stimulus genres (single vs mixed) and number of the features to explore where the differences in the model prediction rates might lie. For this purpose, we grouped the studies into small single genre/multigenre studies, medium single genre/multigenre studies, and medium-large single genre/multigenre studies, and huge multigenre studies. The small single genre/multigenre studies generally performed best (_r_ = 0.836, 95% CI: -0.197-0.989), followed by huge multi-genre studies (_r_ = 0.720, 95% CI: -0.189-0.965) while the medium and medium-large sized studies performance was between close the overall average (medium-large, _r_ = 0.669, 95% CI: 0.347-0.850 and medium _r_ = 0.587, 95% CI: 0.433-0.708). The heterogeneity was lowest in the huge multigenre studies ($\tau^2$ = 0.013) and highest in the small single genre/multigenre studies ($\tau^2$ = 0.318).

@@

>>>>

---
title             : "A Meta-Analysis of Music Emotion Recognition Studies"
shorttitle        : "Meta-analysis of music emotion recognition"

@@

The initial search yielded 553 potential studies after excluding duplicate entries. We interactively screened them for relevance in three stages, resulting in 46 studies that passed our inclusion criteria (music emotion studies using classification or regression methods to predict emotion ratings of music using symbolic or audio features, and containing sufficient detail to convert results to $r$ or $MCC$ values see [SI](http://) for a breakdown). After the screening stage, we defined a set of entities to extract characterising (i) music (genre, stimulus number \[N\], duration), (ii) features extracted (number, type, source, defined by [@panda2020audio]), (iii) model type (regression, neural network, SVM, etc.) and outcome measure ($R^2$, *MSE*, *MCC*), (iv) model complexity (i.e., approximate number of features used to predict ratings), and (v) type of model cross-validation.

We converted all regression results from $R^2$ values into $r$ values for valence and arousal, and classification results into Matthews correlation coefficient [*MCC*, @chicco2020advantages]. To increase consistency in our analyses, we excluded studies using incompatible features (e.g., spectrograms of audio files [@nag2022]), or dependent variables (e.g., one regression study analyzed valence and arousal together, but not separately [@chin2018]).

@@

The search yielded studies of variable (and occasionally questionable) quality. To mitigate potentially spurious effects resulting from the inclusion of low-quality studies, [we excluded studies lacking sufficient details about stimuli, analyzed features, or model architecture](include%20approx%20#%20for%20each?%20This%20would%20be%20good%20since%20the%20rest%20of%20this%20para%20already%20mentions%20counts.%20If%20this%20take%20a%20long%20time,%20we%20could%20save%20this%20for%20an%20additional%20analysis). Finally, we excluded studies published in journals of questionable relevance/quality, (e.g., *Mathematical Problems in Engineering* ceased publication following 17 retractions published between July and September 2024). Overall this step eliminated 12 studies, leaving us with 34 studies in total.

@@

TABLE 1: Summary of data (part of `analysis/preprocessing.qmd` updated 21/1/2025)

|            | Regression                      | Classification            | Total |
|:-----------|:--------------------------------|:--------------------------|:------|
| Study N    | 22                              | 12                        | 34    |
| Model N    | 204                             | 86                        | 290   |
| Techniques | Neural Nets (NN): 66            | 25                        | 91    |
| Techniques | Flexible Discriminants (FD): 50 | 26                        | 76    |
| Techniques | Linear Methods (LM): 62         | 13                        | 75    |
| Techniques | Random Forests (RF): 14         | 10                        | 34    |
| Techniques | KS, Add. & KNN (KS): 12         | 12                        | 24    |
| Feature N  | Min=3, Md=605, Max=6670         | Min=8, Md=126, Max=8904   | NA    |
| Stimulus N | Min=20, Md=324, Max=2486        | Min=124, Md=300, Max=5192 | NA    |

Although the total number of studies meeting the criteria described in the previous section is modest (34 in total), they encompass a large array of models (290 in total) with a relatively even distribution among the three most popular techniques: flexible discriminants, neural nets, and linear methods. The number of features and stimuli within these studies varies significantly, ranging from as few as three features [@battcock2021in] to a maximum of almost 9000 features [@zhang2023mo]. The median number of features differs between regression (605) and classification (126) studies, primarily reflecting the nature of the datasets used in each approach. The number of stimuli is typically around 300-400 (with a median of 324 for regression and 300 for classification), though there is substantial variation, with the extremes from 20 stimuli in @beveridge2018po to 5192 stimuli in @alvarez2023ri. There are also additional dimensions to consider, such as the type of cross-validation used, the music genres analyzed (whether a single genre, multiple genres, or a mix), the type of journal in which the studies were published, and the extraction tool used to extract features. However, these variables do not lend themselves to a simple summary, so we will revisit them during the interpretation and discussion stages.

@@

See `analysis/analysis.qmd` (updated 17/1/2025)

@@

| Concept      | Models, obs | $r$ \[95%-CI\]          | $t$   | $p$   | $\tau^2$ | $I^2$ |
|:-------------|:------------|:------------------------|:------|:------|:---------|:------|
| Valence All  | 102,60017   | 0.583 \[0.541-0.623\]   | 21.40 | .0001 | 0.094    | 97.7% |
| Valence Max  | 22,14172    | 0.669 \[0.560-0.755\]   | 9.58  | .0001 | 0.148    | 98.4% |
| *N Features* |             |                         |       |       |          |       |
| \<18 F       | 5,3036      | 0.811 \[0.542-0.823\]   | \-    | \-    | 0.182    | 98.9% |
| 18-260 F     | 11,7042     | 0.584 \[0.534-0.931\]   | \-    | \-    | 0.133    | 97.6% |
| 260+ F       | 5,3494      | 0.710 \[0.542-0.823\]   | \-    | \-    | 0.044    | 97.2% |
| *Techniques* |             |                         |       |       |          |       |
| KS           | 1,1838      | 0.466 \[-0.634-0.942\]  | \-    | \-    | 0.0185   | 95.1% |
| LM           | 8,1762      | 0.797 \[0.614-0.899\]   | \-    | \-    | 0.1370   | 96.4% |
| FD           | 6,4993      | 0.656 \[0.484-0.779\]   | \-    | \-     | 0.0574  | 96.9% |
| NN           | 4,2249      | 0.393 \[-0.355-0.835\]  | \-    | \-     | 0.0761  | 97.1% |
| RF           | 3,3330      | 0.750 \[0.292-0.928\]   | \-    | \-     | 0.093   | 98.5% |

<!-- table updated TE 21/1/2025 -->

The results indicate that valence can generally be predicted with moderately accuracy, with the best model from each of the 22 studies achieving an average correlation of _r_ = 0.669 (95% CI: 0.560-0.755), called "valence Max" in Table 2. However, when considering all models across these studies (n = 102), the overall prediction rate drops significantly to _r_ = 0.583. We argue that this lower correlation is likely due to the inclusion of baseline models reported in these studies, which may not reflect the true success of the task for the purposes of our analysis.

Further analysis of between-study heterogeneity, as indexed by the $\tau^2$ (0.148) and Higgins & Thompson's $I^2$ statistic [@higgins2002quantifying] at 98.4%, reveals substantial heterogeneity. Since $I^2$ is heavily influenced by study size (with larger N leading to lower sampling error), its value may be less insightful in this context. In contrast, $\tau^2$, which is less sensitive to the number of studies and directly linked to the outcome metric (_r_), provides a more reliable measure of heterogeneity in this case. Also, we note that because the overall heterogeneity in the data is high, we are cautious in our interpretation of the publication bias [@van-aertwicherts_2016].

To better understand the effects across studies and the nature of the observed heterogeneity, Figure 2 presents (A) a forest and (B) funnel plot of the random-effects model, based on the best-performing models from all studies. In terms of the forest plot, the range of prediction values (correlations) is broad, spanning from 0.13 to 0.92, with all studies except Koh et al. (2023) demonstrating evidence of positive correlations. A mean estimate of 0.67 is achieved by 15 out of the 22 models. While the confidence intervals are generally narrow due to the large sample sizes in each study, there are exceptions, such as smaller sample sizes in @beveridge2018po (n = 20), and in @griffiths2021am (n = 40). The funnel plot in panel B of Figure 2 shows clustering at the top of the plot (studies with low standard error) and no assumed larger diversity in the correlations when the error rates increase. However, there is no clear asymmetry in the plot, verified by non-significant Egger's test ($\beta$=5.05, CI95% -0.99-11.09, _t_ = 1.64, _p_ = 0.112, @eggersmith_1997). Coming back to the mean of valence correlation of 0.67 by all studies and the possible impact of study heterogeneity on this estimation, we also calculated the correlation without the studies that lie outside the 95% CI for pooled effect. This left 12 studies in the data and resulted in the meta-analytical pooled correlation of 0.686 (CI95% 0.635-0.731). In other words, despite the large variation in the correlations and standard errors across the studies, this variation in itself does not seem to be a significant driver behind the overall effect.

<!-- numbers in the text updated up to here, TE 21/1/2025 -->

@@

To gain insights into the factors contributing to the wide range of model success, we explored several ways of splitting the data. Table 2 presents two key splits: one based on the number of features used, which we hypothesized might influence model performance, and another based on the modeling techniques employed. In terms of feature sets, we categorized them into three groups: few features (<18), a large number of features (18–260), and massive feature sets (260+). Interestingly, models (4 in total) using a relatively small number of features (<18) performed best (_r_ = 0.817, 95% CI: 0.361–0.958) compared to those utilizing larger feature sets. However, it is worth noting that the models using massive feature sets (260+, 7 studies in total) also performed well (_r_ = 0.702), achieving better and more consistent results than the overall prediction rate (_r_ = 0.669). This observation is supported by the lowest heterogeneity index for the massive feature set group ($\tau^2$ = 0.029), indicating more consistent results across studies. Studies with large number of features (18-260 features, 10 studies in total) delivered the worst results, _r_ = 0.586 (95% CI: 0.360–0.748). Despite the fluctuation in the overall model accuracy between the number of features, the differences are not substantially large to pass the test of statistical significance (_Q_(2) = 3.47, _p_=.176).

<!-- numbers in the text updated up to here, TE 21/1/2025 -->

When analyzing the studies across the five modeling techniques used, the prediction do differ significantly (_Q_(4) = 45.7, _p_ < .0001). Notably, linear models (LM) and flexible discriminants (FD) were the most common, with 8 and 6 studies, respectively, allowing for more confident interpretations. Linear models achieved the highest prediction rate (_r_ = 0.784, 95% CI: 0.625–0.881), though this may be influenced by the smaller datasets typically used in these studies. These studies also exhibited substantially higher heterogeneity ($\tau^2$ = 0.137) compared to other techniques, where $\tau^2$ values were less than half of this. Flexible discriminants performed at a similar level to the overall model average (_r_ = 0.656, 95% CI: 0.484–0.779). While there were only 3 studies involving random forests, these performed well, achieving _r_ = 0.750, 95% CI: 0.292–0.928), and the relatively poor performance of the neural network models represented in four studies (_r_ = 0.340, 95% CI: -0.097–0.668) is difficult to explain without a deeper examination of the specific model architectures and the stimuli used in these studies. Kernel smoothing and KNN models (with only 1 study represented) achieved relatively poor success as well, _r_ = 0.383, 95% CI: 0.348–0.426).

We also ran the sub-grouping analyses across a combination of stimulus genres (single vs mixed) and number of the features to explore where the differences in the model prediction rates might lie. For this purpose, we grouped the studies into small single genre/multigenre studies, medium single genre/multigenre studies, and medium-large single genre/multigenre studies, and huge multigenre studies. The small single genre/multigenre studies generally performed best (_r_ = 0.836), followed by huge multi-genre studies (_r_ = 0.693) while the medium and medium-large sized studies performance was between close the overall average (medium-large, _r_ = 0.768 and medium _r_ = 0.607). The heterogeneity was lowest in the huge multigenre studies ($\tau^2$ = 0.013) and highest in the small single genre/multigenre studies ($\tau^2$ = 0.318).

@@

<!-- numbers in the text updated up to here, TE 21/1/2025 -->
